### 2022-11-07 abstract.tex
Adds devcontainer and magic comments

\chapter{Introduction}

In \todo{Chapter design}, we consider approximately optimal Bayesian experimental design
using an adaptive row sampling algorithm based on DPPs and show that it provides
good approximations. Through generalizing the proof techniques \todo{Determinant preserving random matrices}
applied here, in \todo{Chapter double descent} an extension of the DPP-based design is analyzed in
closed-form for the over-parameterized $n < d$ regime and predicts a double-descent phenomenon in
linear regression which closely matches empirical experiments.
In \todo{Chapter random projections} we isolate the part of the proof involving concentration
of bilinear forms of matrix resolvents away from the DPP-based design in order to obtain bounds on
expected projections $X_S^\dag X_S$ when $X_S = S X$ is obtained by sub-Gaussian sketching matrix $S$.

The second three chapters concern the setting of probabilistic programming, a computational tool
for specifying probabilistic models and automating inference through MCMC and VI. In particular, we  
study approximating a target density $p(x)$ with approximations $q(x)$ for the purposes of
importance sampling and variational inference.
In \todo{Chapter LIC}, we consider parameterizing $q(x)$ by graph neural networks which condition each
node on its Markov blanket. This reduces the conditioning sets for a node, resulting in improvements over
\cite{le2017inference} and run-times which depend on sparsity in the graphical model rather than
the length of execution traces.
When the target $p(x)$ is both multivariate and heavy tailed, \todo{Chapter FTVI}
considers the problem of tail anisotropy through both a theoretical and practical perspective.
We establish that prior fat-tailed estimators \citep{jaini2020tails} are tail isotropic, propose
an anisotropic approximation (fat-tailed variational inference, FTVI) where
an anisotropic product base measures is pushed forwards through a bijective neural network,
and confirm that in practice FTVI improves both density estimation as well as variational inference.
However, we find in practice FTVI and other tail-adaptive approximations often have trouble optimizing
the tail parameter. In \todo{Chapter HTA}, we consider addressing this issue during static analysis of a
probabilistic program's source code. We define generalized Gamma tail asymptotics for a number of
elementary distributions and establish how the tail asymptotics are transformed under algebraic transformations
such as sums and products. This enables a priori computation of tail parameters, which we show
improves the stability and convergence of a number of inference tasks.

>>>>

% !TeX root = thesis.tex
% (This file is included by thesis.tex; you do not latex it by itself.)

\begin{abstract}

    % The text of the abstract goes here.  If you need to use a \section
    % command you will need to use \section*, \subsection*, etc. so that
    % you don't get any numbering.  You probably won't be using any of
    % these commands in the abstract anyway.

    In \cref{ch:design}, we consider approximately optimal Bayesian experimental design
    using an adaptive row sampling algorithm based on DPPs and show that it provides
    good approximations. Through generalizing the proof techniques \todo{Determinant preserving random matrices}
    applied here, in \cref{ch:double_descent} an extension of the DPP-based design is analyzed in
    closed-form for the over-parameterized $n < d$ regime and predicts a double-descent phenomenon in
    linear regression which closely matches empirical experiments.
    In \cref{ch:projections} we isolate the part of the proof involving concentration
    of bilinear forms of matrix resolvents away from the DPP-based design in order to obtain bounds on
    expected projections $X_S^\dag X_S$ when $X_S = S X$ is obtained by sub-Gaussian sketching matrix $S$.

    The second three chapters concern the setting of probabilistic programming, a computational tool
    for specifying probabilistic models and automating inference through MCMC and VI. In particular, we
    study approximating a target density $p(x)$ with approximations $q(x)$ for the purposes of
    importance sampling and variational inference.
    In \cref{ch:lic}, we consider parameterizing $q(x)$ by graph neural networks which condition each
    node on its Markov blanket. This reduces the conditioning sets for a node, resulting in improvements over
    \cite{le2017inference} and run-times which depend on sparsity in the graphical model rather than
    the length of execution traces.
    When the target $p(x)$ is both multivariate and heavy tailed, \cref{ch:ftvi}
    considers the problem of tail anisotropy through both a theoretical and practical perspective.
    We establish that prior fat-tailed estimators \citep{jaini2020tails} are tail isotropic, propose
    an anisotropic approximation (fat-tailed variational inference, FTVI) where
    an anisotropic product base measures is pushed forwards through a bijective neural network,
    and confirm that in practice FTVI improves both density estimation as well as variational inference.
    However, we find in practice FTVI and other tail-adaptive approximations often have trouble optimizing
    the tail parameter. In \cref{ch:hta}, we consider addressing this issue during static analysis of a
    probabilistic program's source code. We define generalized Gamma tail asymptotics for a number of
    elementary distributions and establish how the tail asymptotics are transformed under algebraic transformations
    such as sums and products. This enables a priori computation of tail parameters, which we show
    improves the stability and convergence of a number of inference tasks.

\end{abstract}

### 2022-11-07 intro.tex
Adds devcontainer and magic comments

\chapter{Introduction}

In \todo{Chapter design}, we consider approximately optimal Bayesian experimental design
using an adaptive row sampling algorithm based on DPPs and show that it provides
good approximations. Through generalizing the proof techniques \todo{Determinant preserving random matrices}
applied here, in \todo{Chapter double descent} an extension of the DPP-based design is analyzed in
closed-form for the over-parameterized $n < d$ regime and predicts a double-descent phenomenon in
linear regression which closely matches empirical experiments.
In \todo{Chapter random projections} we isolate the part of the proof involving concentration
of bilinear forms of matrix resolvents away from the DPP-based design in order to obtain bounds on
expected projections $X_S^\dag X_S$ when $X_S = S X$ is obtained by sub-Gaussian sketching matrix $S$.

The second three chapters concern the setting of probabilistic programming, a computational tool
for specifying probabilistic models and automating inference through MCMC and VI. In particular, we  
study approximating a target density $p(x)$ with approximations $q(x)$ for the purposes of
importance sampling and variational inference.
In \todo{Chapter LIC}, we consider parameterizing $q(x)$ by graph neural networks which condition each
node on its Markov blanket. This reduces the conditioning sets for a node, resulting in improvements over
\cite{le2017inference} and run-times which depend on sparsity in the graphical model rather than
the length of execution traces.
When the target $p(x)$ is both multivariate and heavy tailed, \todo{Chapter FTVI}
considers the problem of tail anisotropy through both a theoretical and practical perspective.
We establish that prior fat-tailed estimators \citep{jaini2020tails} are tail isotropic, propose
an anisotropic approximation (fat-tailed variational inference, FTVI) where
an anisotropic product base measures is pushed forwards through a bijective neural network,
and confirm that in practice FTVI improves both density estimation as well as variational inference.
However, we find in practice FTVI and other tail-adaptive approximations often have trouble optimizing
the tail parameter. In \todo{Chapter HTA}, we consider addressing this issue during static analysis of a
probabilistic program's source code. We define generalized Gamma tail asymptotics for a number of
elementary distributions and establish how the tail asymptotics are transformed under algebraic transformations
such as sums and products. This enables a priori computation of tail parameters, which we show
improves the stability and convergence of a number of inference tasks.

>>>>

% !TeX root = thesis.tex
\chapter{Introduction}

In \cref{ch:design}, we consider approximately optimal Bayesian experimental design
using an adaptive row sampling algorithm based on DPPs and show that it provides
good approximations. Through generalizing the proof techniques \todo{Determinant preserving random matrices}
applied here, in \cref{ch:double_descent} an extension of the DPP-based design is analyzed in
closed-form for the over-parameterized $n < d$ regime and predicts a double-descent phenomenon in
linear regression which closely matches empirical experiments.
In \cref{ch:projections} we isolate the part of the proof involving concentration
of bilinear forms of matrix resolvents away from the DPP-based design in order to obtain bounds on
expected projections $X_S^\dag X_S$ when $X_S = S X$ is obtained by sub-Gaussian sketching matrix $S$.

The second three chapters concern the setting of probabilistic programming, a computational tool
for specifying probabilistic models and automating inference through MCMC and VI. In particular, we
study approximating a target density $p(x)$ with approximations $q(x)$ for the purposes of
importance sampling and variational inference.
In \cref{ch:lic}, we consider parameterizing $q(x)$ by graph neural networks which condition each
node on its Markov blanket. This reduces the conditioning sets for a node, resulting in improvements over
\cite{le2017inference} and run-times which depend on sparsity in the graphical model rather than
the length of execution traces.
When the target $p(x)$ is both multivariate and heavy tailed, \cref{ch:ftvi}
considers the problem of tail anisotropy through both a theoretical and practical perspective.
We establish that prior fat-tailed estimators \citep{jaini2020tails} are tail isotropic, propose
an anisotropic approximation (fat-tailed variational inference, FTVI) where
an anisotropic product base measures is pushed forwards through a bijective neural network,
and confirm that in practice FTVI improves both density estimation as well as variational inference.
However, we find in practice FTVI and other tail-adaptive approximations often have trouble optimizing
the tail parameter. In \cref{ch:hta}, we consider addressing this issue during static analysis of a
probabilistic program's source code. We define generalized Gamma tail asymptotics for a number of
elementary distributions and establish how the tail asymptotics are transformed under algebraic transformations
such as sums and products. This enables a priori computation of tail parameters, which we show
improves the stability and convergence of a number of inference tasks.
